---
title: "AF Impact Report subject line testing"
output:
  html_notebook:
    code_folding: hide
---

# Libraries

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(splines)
```

# Import data

```{r}
filepath <- '2018-12-07 Impact Report open click data.xlsx'

recipients <- read_xlsx(
  path = filepath
  , sheet = 'With'
) %>% mutate_if(
  # Replace 0
  is.numeric, replace_na, replace = 0
) %>% mutate(
  # Column data types
  FIRST_KSM_YEAR = as.numeric(FIRST_KSM_YEAR)
  , PROGRAM_GROUP = factor(PROGRAM_GROUP) %>% relevel(ref = 'FT')
  , KLC_CURRENT = !is.na(KLC_CURRENT)
  , KLC_LYBUNT = !is.na(KLC_LYBUNT)
  , AF_STATUS_FY_START = factor(AF_STATUS_FY_START) %>% relevel(ref = 'Lapsed/Non')
  , n = 1
) %>% mutate(
  # Derived variables
  # Any opens or clicks
  DID_OPEN = ifelse(OPENS > 0, TRUE, FALSE)
  , DID_CLICK = ifelse(IMPACT_REPORT_CLICKS > 0, TRUE, FALSE)
  # Decade
  , DECADE = {FIRST_KSM_YEAR/10} %>% floor()
    , DECADE = DECADE * 10
  # AF segment
  , AF_SEG = case_when(
    str_detect(AF_GIVING_SEGMENT, 'Loyal 3+') ~ 'Loyal3+'
    , str_detect(AF_GIVING_SEGMENT, 'Loyal') ~ 'Loyal2'
    , TRUE ~ AF_STATUS
  ) %>% factor()
    %>% relevel(ref = 'Lapsed/Non')
  # Subject line factors
  , HAS_KSM = if_else(str_detect(str_to_lower(EMAIL_SUBJECT), 'kellogg'), TRUE, FALSE)
  , HAS_YOUR = ifelse(str_detect(str_to_lower(EMAIL_SUBJECT), 'your'), TRUE, FALSE)
  , IS_SHORT = ifelse(str_count(EMAIL_SUBJECT, ' ') < 3, TRUE, FALSE)
)
```

# Exploration

Check that I categorized the subject lines correctly.

```{r}
# Check the email factors were correctly assigned
recipients %>% group_by(EMAIL_SUBJECT) %>%
  summarise(`Has KSM` = max(HAS_KSM), `Has your` = max(HAS_YOUR), `Is short` = max(IS_SHORT))
```

Good.

Check that there are no systematic differences between which groups received each subject line. (This was already done via block randomization but is verified below.)

```{r}
# Subject line by program
recipients %>% xtabs(n ~ EMAIL_SUBJECT + PROGRAM_GROUP, data = .)
```

Good.

```{r}
# Subject line by decade
recipients %>% xtabs(n ~ EMAIL_SUBJECT + DECADE, data = .)
```

Still good.

```{r}
# Subject line by AF segment
recipients %>% xtabs(n ~ EMAIL_SUBJECT + AF_SEG, data = .)
```

Still good.

Look at the summary statistics for each subject line.

```{r}
# Overall stats  by subject line
recipients %>% group_by(EMAIL_SUBJECT) %>%
  summarise(
    recipients = length(MSG_ID)
    , `total opens` = sum(OPENS)
    , `total clicks` = sum(TOTAL_CLICKS)
    , `open rate` = mean(DID_OPEN) %>% scales::percent()
    , `click rate` = mean(DID_CLICK) %>% scales::percent()
  ) %>% rbind(
    c(
      EMAIL_SUBJECT = 'TOTAL'
      , recipients = nrow(recipients)
      , `total opens` = sum(recipients$OPENS)
      , `total clicks` = sum(recipients$TOTAL_CLICKS)
      , `open rate` = mean(recipients$DID_OPEN) %>% scales::percent()
      , `click rate` = mean(recipients$DID_CLICK) %>% scales::percent()
    )
  ) %>% rename(
    `email subject` = EMAIL_SUBJECT
  )
```

Note that total opens and total clicks may count multiple interactions from a single individual, but the rates are deduped. Eyeballing this, it looks like including "Kellogg" decreases open rates but this needs to be tested.

# Exploratory classification model

Consider a standard binary logistic regression model, where the response variable is DID_OPEN (i.e. people who didn't open the email are coded as 0, and people who opened the email 1+ times are coded as 1), and we're interested in the effect of the three factors of interest (short versus long, Kellogg versus no Kellogg, your versus no your) after controlling for the various demographic variables.

```{r}
glmod <- glm(
  DID_OPEN ~
    HAS_KSM * HAS_YOUR * IS_SHORT
    + AF_SEG + AF_STATUS_FY_START + KLC_LYBUNT + KLC_CURRENT
    + PROGRAM_GROUP + ns(FIRST_KSM_YEAR, df = 5)
  , data = recipients
  , family = binomial
)
```

```{r}
summary(glmod)
```

The following appears to be the case:

  * Including "Kellogg" in the subject line reduces the open rate on average
  * Including "Your" in the subject line reduces the open rate on average
  * There is no main effect associated with short versus long subject lines
  * There is no evidence for two-way or three-way interaction effects on the open rate
  * KLC status (current and previous year) is associated with increased open rate
  * First KSM year is nonlinearly associated with open rate
  
I doubt there is enough data to see much of interest with clicks.

```{r}
glmod %>% update(DID_CLICK ~ ., data = recipients) %>% summary()
```

  * Including "Your" is also associated with fewer clicks
  * Last year's KLC donors are somewhat more likely to click
  * Part-time alumni are somewhat less likely to click

Next, consider the model without interaction effects.

```{r}
glmod %>%
  update(. ~ . - HAS_KSM * HAS_YOUR * IS_SHORT + HAS_KSM + HAS_YOUR + IS_SHORT, data = recipients) %>%
  summary()
```

This is essentially the same AIC as the model with interactions.

Finally, consider a reduced model without 3-way interactions and the non-impactful variables.

```{r}
glmod_reduced <- glm(
  DID_OPEN ~
    HAS_KSM * HAS_YOUR * IS_SHORT - HAS_KSM:HAS_YOUR:IS_SHORT
    + KLC_LYBUNT + KLC_CURRENT
    + PROGRAM_GROUP + ns(FIRST_KSM_YEAR, df = 5)
  , data = recipients
  , family = binomial
)
```
```{r}
summary(glmod_reduced)
```

Message length is significant only when the interaction terms are excluded.

# Model calibration results

```{r}
data.frame(
    class = recipients$DID_OPEN + 0
  , prediction = predict(glmod, type = 'response')
) %>%
# Plot results
ggplot(aes(x = prediction, y = class)) +
  geom_point(aes(color = as.factor(class))) +
  geom_smooth(method = 'loess') +
  geom_abline(slope = 1, intercept = 0) +
  labs(title = 'Loess smoother on predicted open rate', color = 'class') +
  xlim(c(0,1))
```

That's an outstanding result.

```{r}
data.frame(
    class = recipients$DID_OPEN + 0
  , prediction = predict(glmod_reduced, type = 'response')
) %>%
# Plot results
ggplot(aes(x = prediction, y = class)) +
  geom_point(aes(color = as.factor(class))) +
  geom_smooth(method = 'loess') +
  geom_abline(slope = 1, intercept = 0) +
  labs(title = 'Loess smoother on predicted open rate, reduced model', color = 'class') +
  xlim(c(0,1))
```

Still an excellent result with the reduced model.

# Class year effects

```{r}
data.frame(
  FIRST_KSM_YEAR = recipients$FIRST_KSM_YEAR
  , y.partial.resid = residuals(glmod, type = 'partial')[, 'ns(FIRST_KSM_YEAR, df = 5)']
  , class = as.factor(recipients$DID_OPEN + 0)
) %>%
  ggplot(aes(x = FIRST_KSM_YEAR, y = y.partial.resid)) +
  geom_point(aes(color = class, fill = class), alpha = .1, size = 1) +
  geom_hline(yintercept = 0, color = 'darkgray') +
  geom_smooth(method = 'loess') +
  scale_x_continuous(breaks = seq(0, 3000, by = 10)) +
  scale_y_continuous(breaks = seq(-10, 10, by = 2)) +
  labs(title = 'First KSM year spline versus partial residuals for email open classification'
       , x = 'First KSM year'
       , y = 'partial residuals')
```

This is exactly as expected -- all else equal, younger alums are more likely to open the email than older ones.
