---
title: "AF Impact Report randomization"
output: html_notebook
---

# Libraries

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(randomizr)
```


# Import data

```{r}
filepath <- 'Annual Fund Impact Report Email List.xlsx'

participation <- read_xlsx(
  path = filepath
  , sheet = 'Participation'
)

klc <- read_xlsx(
  path = filepath
  , sheet = 'KLC'
)

data <- rbind(participation, klc) %>%
  select(ID, PROG1, YEAR1)
remove(participation, klc)
```

# Transform data

```{r}
data <- data %>%
  mutate(
    DECADE = {as.numeric(YEAR1)/10} %>% floor()
    , DECADE = DECADE * 10
    , PRG = str_extract(PROG1, pattern = '.*(?=-)')
    , PRG = ifelse(is.na(PRG), PROG1, PRG) %>% factor()
    , n = 1
  )
```

# Exploration

```{r}
data %>% xtabs(n ~ DECADE + PRG, data = .)
```

Hypotheses:

  * Older alumni interact with email differently than younger ones.
  * FT alumni interact with KSM emails differently than others.

  - One block could be 1930-1960, all programs
  - Another block for 1970
  - Two blocks for the other decades: FT, and everyone else
  - 10 blocks total

# Block randomization

```{r}
data <- data %>% mutate(
  block = case_when(
    DECADE <= 1960 ~ 1
    , DECADE == 1970 ~ 2
    , DECADE == 1980 & PRG == 'FT' ~ 3
    , DECADE == 1980 ~ 4
    , DECADE == 1990 & PRG == 'FT' ~ 5
    , DECADE == 1990 ~ 6
    , DECADE == 2000 & PRG == 'FT' ~ 7
    , DECADE == 2000 ~ 8
    , DECADE == 2010 & PRG == 'FT' ~ 9
    , DECADE == 2010 ~ 10
  )
)
```

```{r}
data %>% xtabs(n ~ DECADE + PRG + block, data = .)
```

Looks fine.

# Block randomization

```{r}
set.seed(1001931)
conds <- as.character(1:8)
data$randomized <- with(data
  , block_ra(block, conditions = conds)
)
```

Check results.

```{r}
data %>% xtabs(n ~ block + randomized, data = .)
```

# Export

```{r}
write.csv(data, file = 'randomized.csv', row.names = FALSE)
```

