---
title: "01 DCA Giving Trends"
output:
  html_notebook:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    code_folding: hide
---

# Goal

Examine giving trends for the Dean's Consulting Alliance Spring 2019 group. Specifically, assess the following hypotheses:

  * If a graduate disengages with Kellogg right after graduation, they don’t donate ever
  * Mature alumni who donate in later career have donated as a recent grad
  
# Data

A [SQL query](https://github.com/phively/ksm-stats/blob/master/2019-DCA-recent-grad-giving/code/sql-giving-query.sql) was run to pull all full-time alumni (defined as program in 2Y, 1Y, MMM, JDMBA, FT) along with their program, master's degree year, fiscal year of first gift, giving in FY13-18, and giving the fiscal year of graduation and in each of the five subsequent fiscal years.

```{r setup, message = FALSE, warning = FALSE}
# R libraries
library(tidyverse)
library(ggplot2)
library(readxl)
```

```{r}
# Import data
giving <- read_xlsx(
  path = 'data/2019-04-29 giving data.xlsx'
  , sheet = 'With'
)

# Cleanup
giving <- giving %>% filter(
  # Drop 0 years
  FIRST_MASTERS_YEAR != '0000'
) %>% mutate(
  # Program as factor
  PROGRAM = factor(PROGRAM)
  # Year as numeric
  , FIRST_MASTERS_YEAR = as.numeric(FIRST_MASTERS_YEAR)
) %>% mutate_at(
  # Convert yearly giving totals into a 0-1 indicator column
  .vars = vars(GIVING_PFY1:GIVING_GRADYR5)
  , .funs = function(x) {
    ifelse(replace_na(x, 0) == 0, 'nondonor', 'donor') %>%
      factor() %>%
      relevel(ref = 'nondonor')
  }
)
```

## Distribution by program and year

```{r, fig.height = 7, fig.width = 12}
giving %>% 
  ggplot(aes(x = FIRST_MASTERS_YEAR, color = PROGRAM, fill = PROGRAM)) +
  geom_bar(aes(y = ..count..)) +
  facet_grid(PROGRAM ~ .) +
  scale_x_continuous(breaks = seq(0, 3000, by = 5), minor_breaks = 0:3000) +
  scale_y_continuous(breaks = seq(0, 1E5, by = 100)) +
  theme(
    panel.grid.major = element_line(color = 'darkgray')
    , panel.grid.minor = element_line(color = NA)
  )
```

## Donor status by year

```{r, fig.width = 12, fig.height = 4}
giving %>% 
  ggplot(aes(x = FIRST_MASTERS_YEAR, color = GIVING_PFY1, fill = GIVING_PFY1)) +
  geom_bar(aes(y = ..count..), position = 'stack') +
  scale_x_continuous(breaks = seq(0, 3000, by = 5), minor_breaks = 0:3000) +
  scale_y_continuous(breaks = seq(0, 1E5, by = 100)) +
  theme(
    panel.grid.major = element_line(color = 'darkgray')
    , panel.grid.minor = element_line(color = NA)
  ) +
  labs(title = paste('FY', min(giving$CURR_FY - 1, na.rm = TRUE), 'donor status by MBA year'))
```

