---
title: "01 DCA Giving Trends"
output:
  html_notebook:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    code_folding: hide
---

# Goal

Examine giving trends for the Dean's Consulting Alliance Spring 2019 group. Specifically, assess the following hypotheses:

  * If a graduate disengages with Kellogg right after graduation, they don’t donate ever
  * Mature alumni who donate in later career have donated as a recent grad
  
# Data

A [SQL query](https://github.com/phively/ksm-stats/blob/master/2019-DCA-recent-grad-giving/code/sql-giving-query.sql) was run to pull all full-time alumni (defined as program in 2Y, 1Y, MMM, JDMBA, FT) along with their program, master's degree year, fiscal year of first gift, giving in FY13-18, and giving the fiscal year of graduation and in each of the five subsequent fiscal years.

```{r setup, message = FALSE, warning = FALSE}
# R libraries
library(tidyverse)
library(ggplot2)
library(readxl)
```

```{r}
# Import data
giving <- read_xlsx(
  path = 'data/2019-04-29 giving data.xlsx'
  , sheet = 'With'
)

# Cleanup
giving <- giving %>% filter(
  # Drop 0 years
  FIRST_MASTERS_YEAR != '0000'
) %>% mutate(
  # Program as factor
  PROGRAM = factor(PROGRAM)
  # Year as numeric
  , FIRST_MASTERS_YEAR = as.numeric(FIRST_MASTERS_YEAR)
) %>% mutate_at(
  # Convert yearly giving totals into a 0-1 indicator column
  .vars = vars(GIVING_PFY1:GIVING_GRADYR5)
  , .funs = function(x) {
    ifelse(replace_na(x, 0) == 0, 'nondonor', 'donor') %>%
      factor() %>%
      relevel(ref = 'nondonor')
  }
)
```

## Distribution by program and year

```{r, fig.height = 7, fig.width = 12}
giving %>% 
  ggplot(aes(x = FIRST_MASTERS_YEAR, color = PROGRAM, fill = PROGRAM)) +
  geom_bar(aes(y = ..count..)) +
  facet_grid(PROGRAM ~ .) +
  scale_x_continuous(breaks = seq(0, 3000, by = 5), minor_breaks = 0:3000) +
  scale_y_continuous(breaks = seq(0, 1E5, by = 100)) +
  theme(
    panel.grid.major = element_line(color = 'darkgray')
    , panel.grid.minor = element_line(color = NA)
  ) +
  labs(title = 'Full-time alumni counts by class year and program')
```

## Donor status by year

```{r, fig.width = 12, fig.height = 4}
giving %>% 
  ggplot(aes(x = FIRST_MASTERS_YEAR, color = GIVING_PFY1, fill = GIVING_PFY1)) +
  geom_bar(aes(y = ..count..), position = 'stack') +
  scale_x_continuous(breaks = seq(0, 3000, by = 5), minor_breaks = 0:3000) +
  scale_y_continuous(breaks = seq(0, 1E5, by = 100)) +
  theme(
    panel.grid.major = element_line(color = 'darkgray')
    , panel.grid.minor = element_line(color = NA)
  ) +
  labs(title = paste('FY', min(giving$CURR_FY - 1, na.rm = TRUE), 'donor status by MBA year'))
```

# First gift year

Explore the assumption that "if a graduate disengages with Kellogg right after graduation, they don’t donate ever" by looking at the distribution of nondonors and donors, and time from graduation year until first gift year.

Begin by aggregating the data by program, MBA year, and years until the first gift was made.

```{r}
preagg <- giving %>% mutate(
  # Years to first gift
  yrs_to_first_gift = FIRST_GIFT_YEAR - FIRST_MASTERS_YEAR
  # Donor status
  , donor_status = ifelse(is.na(FIRST_GIFT_YEAR), 'nondonor', 'donor') %>%
    factor() %>%
    relevel(ref = 'nondonor')
  # Counter helper
  , n = 1
)

agg <- preagg %>% group_by(
  PROGRAM
  , FIRST_MASTERS_YEAR
  , donor_status
  , yrs_to_first_gift
) %>% summarise(
  n = sum(n)
)

# Save data
agg %>% write_csv(path = 'data/aggregated donor data.csv')
```

Next, examine the overall proportion of nondonors and donors.

```{r}
preagg %>%
  ggplot(aes(x = donor_status, color = donor_status, fill = donor_status)) +
  geom_bar(aes(y = ..count..)) +
  geom_text(aes(label = ..count..)
            , stat = 'count', vjust = 1.5, color = 'black') +
  geom_text(aes(label = scales::percent(..count.. / sum(..count..)), y = ..count..)
            , stat = 'count', vjust = 3, color = 'black') +
  labs(title = 'Full-time donors versus nondonors')
```

More alumni are actually donors than nondonors.

Breaking out the donors by the number of years since graduating it took to make their first gift yields the following.

```{r}
preagg %>%
  filter(donor_status == 'donor') %>%
  ggplot(aes(x = yrs_to_first_gift)) +
  geom_histogram(binwidth = 1, alpha = .5) +
  geom_density(aes( y = ..count..)) +
  scale_x_continuous(breaks = seq(-1E3, 1E3, by = 10)) +
  labs(title = 'Full-time alumni donor years from graduation to first gift')
```

We see that most people who end up donating do so for the first time the year they graduate, and there are minor bumps the year after graduation as well as 5, 10, 15...years after corresponding to the Reunion cycle.
