---
title: "01 DCA Giving Trends"
output:
  html_notebook:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    code_folding: hide
---

# Goal

Examine giving trends for the Dean's Consulting Alliance Spring 2019 group. Specifically, assess the following hypotheses:

  * If a graduate disengages with Kellogg right after graduation, they don’t donate ever
  * Mature alumni who donate in later career have donated as a recent grad
  
# Data

A [SQL query](https://github.com/phively/ksm-stats/blob/master/2019-DCA-recent-grad-giving/code/sql-giving-query.sql) was run to pull all full-time alumni (defined as program in 2Y, 1Y, MMM, JDMBA, FT) along with their program, master's degree year, fiscal year of first gift, giving in FY13-18, and giving the fiscal year of graduation and in each of the five subsequent fiscal years.

```{r setup, message = FALSE, warning = FALSE}
# R libraries
library(tidyverse)
library(ggplot2)
library(readxl)
library(splines)
```

```{r}
# Import data
giving <- read_xlsx(
  path = 'data/2019-04-29 giving data.xlsx'
  , sheet = 'With'
)

# Cleanup
giving <- giving %>% filter(
  # Drop 0 years
  FIRST_MASTERS_YEAR != '0000'
) %>% mutate(
  # Program as factor
  PROGRAM = factor(PROGRAM)
  # Year as numeric
  , FIRST_MASTERS_YEAR = as.numeric(FIRST_MASTERS_YEAR)
) %>% mutate_at(
  # Convert yearly giving totals into a 0-1 indicator column
  .vars = vars(GIVING_PFY1:GIVING_GRADYR5)
  , .funs = function(x) {
    ifelse(replace_na(x, 0) == 0, 'nondonor', 'donor') %>%
      factor() %>%
      relevel(ref = 'nondonor')
  }
)
```

## Distribution by program and year

```{r, fig.height = 7, fig.width = 12}
giving %>% 
  ggplot(aes(x = FIRST_MASTERS_YEAR, color = PROGRAM, fill = PROGRAM)) +
  geom_bar(aes(y = ..count..)) +
  facet_grid(PROGRAM ~ .) +
  scale_x_continuous(breaks = seq(0, 3000, by = 5), minor_breaks = 0:3000) +
  scale_y_continuous(breaks = seq(0, 1E5, by = 100)) +
  theme(
    panel.grid.major = element_line(color = 'darkgray')
    , panel.grid.minor = element_line(color = NA)
  ) +
  labs(title = 'Full-time alumni counts by class year and program')
```

## Donor status by year

```{r, fig.width = 12, fig.height = 4}
giving %>% 
  ggplot(aes(x = FIRST_MASTERS_YEAR, color = GIVING_PFY1, fill = GIVING_PFY1)) +
  geom_bar(aes(y = ..count..), position = 'stack') +
  scale_x_continuous(breaks = seq(0, 3000, by = 5), minor_breaks = 0:3000) +
  scale_y_continuous(breaks = seq(0, 1E5, by = 100)) +
  theme(
    panel.grid.major = element_line(color = 'darkgray')
    , panel.grid.minor = element_line(color = NA)
  ) +
  labs(title = paste('FY', min(giving$CURR_FY - 1, na.rm = TRUE), 'donor status by MBA year'))
```

# First gift year

Explore the assumption that "if a graduate disengages with Kellogg right after graduation, they don’t donate ever" by looking at the distribution of nondonors and donors, and time from graduation year until first gift year.

Begin by aggregating the data by program, MBA year, and years until the first gift was made.

```{r}
preagg <- giving %>% mutate(
  # Years to first gift
  yrs_to_first_gift = FIRST_GIFT_YEAR - FIRST_MASTERS_YEAR
  # Donor status
  , donor_status = ifelse(is.na(FIRST_GIFT_YEAR), 'nondonor', 'donor') %>%
    factor() %>%
    relevel(ref = 'nondonor')
  # Counter helper
  , n = 1
)

agg <- preagg %>% group_by(
  PROGRAM
  , FIRST_MASTERS_YEAR
  , donor_status
  , yrs_to_first_gift
) %>% summarise(
  n = sum(n)
)

# Save data
agg %>% write_csv(path = 'data/aggregated donor data.csv')
```

Next, examine the overall proportion of nondonors and donors.

```{r}
preagg %>%
  ggplot(aes(x = donor_status, color = donor_status, fill = donor_status)) +
  geom_bar(aes(y = ..count..)) +
  geom_text(aes(label = ..count..)
            , stat = 'count', vjust = 1.5, color = 'black') +
  geom_text(aes(label = scales::percent(..count.. / sum(..count..)), y = ..count..)
            , stat = 'count', vjust = 3, color = 'black') +
  labs(title = 'Full-time donors versus nondonors')
```

More alumni are actually donors than nondonors.

Breaking out the donors by the number of years since graduating it took to make their first gift yields the following.

```{r}
preagg %>%
  filter(donor_status == 'donor') %>%
  ggplot(aes(x = yrs_to_first_gift)) +
  geom_histogram(binwidth = 1, alpha = .5) +
  geom_density(aes( y = ..count..)) +
  scale_x_continuous(breaks = seq(-1E3, 1E3, by = 10)) +
  labs(title = 'Full-time alumni donor years from graduation to first gift')
```

We see that most people who end up donating do so for the first time the year they graduate, and there are minor bumps the year after graduation as well as 5, 10, 15...years after corresponding to the Reunion cycle.

# Impact of early giving on later giving

For fundraising purposes, the more relevant question is whether it's true that "mature alumni who donate in later career have donated as a recent grad."

The most straightforward way I can think of to test this hypothesis with the data is to see whether early giving behavior can predict current-day giving after controlling for recent behavior, as well as cohort effects. In other words, if $X$ are the other explanatory variables:

$$ P \left(\text{gave last year} ~ | ~ \text{gave as a recent grad}, X \right) > P \left(\text{gave last year} ~ | ~ !\text{gave as a recent grad}, X \right) $$

Define "gave last year" as making a gift in fiscal year `r max(giving$CURR_FY - 1, na.rm = TRUE)`, "gave as a recent grad" as making a gift 0 through 5 fiscal years after graduation, and recent giving as making a gift in any of fiscal years `r max(giving$CURR_FY - 6, na.rm = TRUE)` through `r max(giving$CURR_FY - 2, na.rm = TRUE)`.

With this in mind, I'll restrict the analysis to only alumni graduating in or before `r max(giving$CURR_FY - 1 - 11, na.rm = TRUE)`, so the full 10 years of giving history are available.

```{r}
glm1 <- giving %>% filter(
  FIRST_MASTERS_YEAR <= CURR_FY - 1 - 11
) %>% select(
  -ID_NUMBER, -REPORT_NAME, -DEGREES_CONCAT, -FIRST_KSM_YEAR, -CURR_FY, -FIRST_GIFT_YEAR
) %>% glm(
  GIVING_PFY1 ~ . - FIRST_MASTERS_YEAR + ns(FIRST_MASTERS_YEAR, df = 3)
  , data = .
  , family = binomial()
)
summary(glm1)
```

This is a binary logistic regression model predicting last year's log odds of making a gift given the predictors. From past experience I know that the effect of graduation year is nonlinear, so I used a natural spline fit with 3 degrees of freedom (arbitrary).

Importantly, we can examine the contrasts between donor/nondonor response for each factor.

```{r}
termplot(glm1)
```

There are large differences between each level for recent giving - GIVING_PFY2 through GIVING_PFY6 (and the coefficients are *very* statistically significant in the regression output), but almost no perceptible difference between the levels for recent alumni giving - GIVING_GRADYR0 through GIVING_GRADYR5 (and these coefficients were not statistically significant, except for the most recent, GIVING_GRADYR5).

But given that the class gift program hasn't existed forever, perhaps there's an interaction between recent alumni giving and graduation year.

```{r}
glm2 <- giving %>% filter(
  FIRST_MASTERS_YEAR <= CURR_FY - 1 - 11
) %>% select(
  -ID_NUMBER, -REPORT_NAME, -DEGREES_CONCAT, -FIRST_KSM_YEAR, -CURR_FY, -FIRST_GIFT_YEAR
) %>% glm(
  GIVING_PFY1 ~ . - FIRST_MASTERS_YEAR + ns(FIRST_MASTERS_YEAR, df = 3)
  + ns(FIRST_MASTERS_YEAR, df = 3):GIVING_GRADYR0
  + ns(FIRST_MASTERS_YEAR, df = 3):GIVING_GRADYR1
  + ns(FIRST_MASTERS_YEAR, df = 3):GIVING_GRADYR2
  + ns(FIRST_MASTERS_YEAR, df = 3):GIVING_GRADYR3
  + ns(FIRST_MASTERS_YEAR, df = 3):GIVING_GRADYR4
  + ns(FIRST_MASTERS_YEAR, df = 3):GIVING_GRADYR5
  , data = .
  , family = binomial()
)

summary(glm2)
```

None of the interactions are significant. I'll try one more approach: what about if I remove all recent giving (not recommended) and see if early giving is predictive?

```{r}
glm3 <- giving %>% filter(
  FIRST_MASTERS_YEAR <= CURR_FY - 1 - 11
) %>% select(
  -ID_NUMBER, -REPORT_NAME, -DEGREES_CONCAT, -FIRST_KSM_YEAR, -CURR_FY, -FIRST_GIFT_YEAR
) %>% glm(
  GIVING_PFY1 ~ . - FIRST_MASTERS_YEAR + ns(FIRST_MASTERS_YEAR, df = 3)
  - GIVING_PFY2 - GIVING_PFY3 - GIVING_PFY4 - GIVING_PFY5 - GIVING_PFY6
  , data = .
  , family = binomial()
)

summary(glm3)
```

We finally have some "significant" coefficients at the cost of blowing up the AIC. I find this pretty conclusive: early giving behavior is only predictive of previous year donor status when recent data is unavailable.