---
title: "Compensation survey"
output:
  html_notebook:
    code_folding: hide
    toc: TRUE
    toc_float:
      collapsed: FALSE
---

<style>
#header .btn-group {
    display: none;
}
</style>

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
# Libraries
library(tidyverse)
library(lubridate)
library(readxl)
library(wranglR)
library(knitr)
library(splines)
library(car)

# Parameters -- edit if needed
data_as_of <- ymd('20191202')

# Set to FALSE to hide all code chunks when rendering the document
opts_chunk$set(echo = FALSE)
```
```{r}
# Load data from last time
load('data/dat_agg4.Rdata')
```

# Goal

Understand the distribution of recent graduate compensation from the exit survey and how it relates to the email survey results.

# Compensation survey data

```{r}
# Load compensation data
dat_comp <- read_xlsx(
  path = 'data/2Y_CL2011-2018_PostMBACompensation.xlsx'
  , sheet = 1
) %>% mutate(
  EmplID = as.numeric(EmplID)
)

# Load emplids
emplids <- read_xlsx(path = 'data/ID to EMPLID conversion.xlsx', sheet = 1) %>%
  mutate(
    ID_NUMBER = as.numeric(CATRACKS_ID)
    , EMPLID = as.numeric(EMPLID)
  ) %>% select(-CATRACKS_ID)

# Load degrees
dat_degs <- read_xlsx(path = 'data/degree lookups.xlsx', sheet = 1)

# Look up IDs and degrees
dat_comp <- dat_comp %>%
  left_join(emplids, by = c('EmplID' = 'EMPLID')) %>%
  left_join(dat_degs %>% select(-ID_NUMBER), by = c('EmplID' = 'EMPLID')) %>%
  # Fix data types
  mutate(
    FIRST_KSM_YEAR = as.numeric(FIRST_KSM_YEAR)
    , FIRST_MASTERS_YEAR = as.numeric(FIRST_MASTERS_YEAR)
    , PROGRAM = as.factor(PROGRAM)
    , PROGRAM_GROUP = as.factor(PROGRAM_GROUP)
    , ClassYear = as.factor(FIRST_MASTERS_YEAR)
    , AnySignon = factor(SignOn > 0)
  )

# Join to dat_agg
dat_agg <- dat_agg %>%
  left_join(
    dat_comp %>% select(ID_NUMBER, BaseSalary, SignOn, Compensation, AnySignon)
    , by = c('ID_NUMBER' = 'ID_NUMBER')
  )
```

First check that `Compensation` = `BaseSalary` + `Signon` and how many map back to survey recipients.

```{r}
dat_comp %>%
  mutate(
    valid_totals = BaseSalary + SignOn == Compensation
    , has_id = !is.na(ID_NUMBER)
  ) %>%
  select(valid_totals, has_id) %>%
  summary()
```

Looks good.

```{r}
dat_comp %>% select(-EmplID, -ID_NUMBER, -FIRST_KSM_YEAR) %>% summary()
```

```{r}
plot_comp_by_yr <- dat_comp %>%
  ggplot(aes(x = BaseSalary, y = SignOn, color = ClassYear)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  geom_rug(position = position_jitter(width = 100, seed = 123)) +
  # Reference line
  geom_abline(slope = -1, intercept = mean(dat_comp$Compensation)) +
  scale_x_continuous(breaks = seq(0, 1E6, by = 50E3), labels = scales::dollar, limits = c(0, NA), expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1E6, by = 20E3), labels = scales::dollar) +
  labs(title = paste0(
   'Base salary and signon bonus by individual response
   (reference line is mean = '
   , mean(dat_comp$Compensation) %>% scales::dollar()
   , ')')
   )

print(plot_comp_by_yr)
```

```{r, fig.height = 12, fig.width = 8}
plot_comp_by_yr +
  facet_grid(ClassYear ~ .)
```

Quite noisy, though there is an association between `BaseSalary` and `SignOn`.

Univariate densities:

```{r}
# Plotting function
plot_univar <- function(data, varname, bw = 10E3) {
  varname <- ensym(varname)
  data %>%
    ggplot(aes(x = !!varname, color = ClassYear)) +
    geom_histogram(aes(fill = ClassYear), alpha = .1, binwidth = bw, boundary = 0) +
    geom_density(aes(y = bw * ..count..)) +
    facet_grid(ClassYear ~ .) +
    scale_x_continuous(labels = scales::dollar, breaks = seq(0, 1E6, by = 50E3)) +
    labs(title = paste('Univariate', varname, 'distribution'))
}
```

```{r, fig.height = 12, fig.width = 8}
dat_comp %>% plot_univar(varname = Compensation)
```

Hypothesis: Compensation is bimodal, with distinct `SignOn > 0` and `SignOn == 0` groups.

```{r, fig.height = 12, fig.width = 8}
dat_comp %>% plot_univar(varname = BaseSalary)
```

Base salary is also bimodal -- probably a result of employer industry.

```{r, fig.height = 12, fig.width = 8}
dat_comp %>% plot_univar(varname = SignOn)
```

```{r}
bw <- 10E3

dat_comp_summary <- dat_comp %>%
  group_by(AnySignon, ClassYear) %>%
  summarise(mean = mean(Compensation))

dat_comp %>%
  ggplot(aes(x = Compensation, color = ClassYear)) +
  geom_density(aes(y = bw * ..count..)) +
  # Reference line
  geom_vline(
    data = dat_comp_summary
    , aes(color = ClassYear, xintercept = mean)
    , linetype = 'dashed'
  ) +
  facet_grid(AnySignon ~ .) +
  scale_x_continuous(labels = scales::dollar, breaks = seq(0, 1E6, by = 50E3)) +
  labs(title = 'Compensation by SignOn', y = 'count')
```

Note that each group is approximately symmetrical. Modeling compensation as a function of the various covariates.

```{r}
lm_comp <- dat_comp %>%
  lm(
    Compensation ~ ClassYear + PROGRAM + AnySignon
    , data = .
  )
```
```{r}
lm_comp %>% summary()
```

```{r}
plot(lm_comp)
```

```{r}
termplot(lm_comp)
```

This works better than I thought it would! It'd be interesting to also look at `SignOn` as a function of e.g. industry but that would require some more data cleanup.

# Email and compensation data

```{r}
dat_agg %>%
  mutate(`Compensation (thousands)` = Compensation/1E3) %>%
  select(`Compensation (thousands)`) %>%
  summary()
```

So only `r {1 - sum(is.na(dat_agg$Compensation)) / nrow(dat_agg)} %>% scales::percent() %>% I()` of the email recipients have salary information available.

```{r, warning = FALSE}
plot_comp_by_pub <- dat_agg %>%
  filter(!is.na(Compensation)) %>%
  ggplot(aes(x = email_group_filled_lb, y = Compensation)) +
  geom_point(position = position_jitter(width = .2, seed = 123), alpha = .1) +
  geom_boxplot(alpha = .5) +
  scale_y_continuous(labels = scales::dollar, breaks = seq(0, 1E6, by = 50E3)) +
  labs(x = 'Email group', title = 'Compensation by Email Group')

# Display results
plot_comp_by_pub
```

```{r, warning = FALSE}
# Break out by SignOn
plot_comp_by_pub +
  facet_grid(. ~ AnySignon) +
  labs(title = 'Compensation by email group and AnySignon')
```

We see the same graduation year and sign-on effect as above.

First look at open rates versus these two variables.

```{r, warning = FALSE, message = FALSE}
plot_opens_comp <- dat_agg %>%
  ggplot(aes(x = Compensation, y = any_open)) +
  geom_point(aes(color = as.factor(any_open))) +
  geom_smooth(method = 'lm', color = 'darkgreen') +
  geom_smooth(method = 'loess') +
  scale_y_continuous(breaks = c(0, 1), minor_breaks = NULL)

# Linear compensation
plot_opens_comp +
  scale_x_continuous(labels = scales::dollar, breaks = seq(0, 1E6, by = 50E3)) +
  labs('Email open rate versus compensation')
```

```{r, message = FALSE, warning = FALSE}
# Log compensation
plot_opens_comp +
  scale_x_continuous(trans = 'log10', labels = scales::dollar, breaks = c(10E3, 50E3, 100E3, 150E3, 200E3, 250E3)) +
  labs('Email open rate versus log10 compensation')
```

Transformation doesn't seem to make that much of a difference. Besides the (low) outliers the linear trend is pretty close.

```{r, warning = FALSE}
dat_agg %>%
  group_by(AnySignon) %>%
  summarise(
    n = n()
    , opens = sum(any_open)
    , open_rate = scales::percent(opens / n)
  )
```

# Emails versus compensation

Repeat the previous email open analysis with compensation and signon as explanatory variables.

```{r}
glm_opens <- dat_agg %>%
  glm(
    any_open ~ email_group_filled + same_email + send_email_status + ns(weeks_since_email, df = 4) + Compensation + AnySignon
    , data = .
    , family = binomial()
  )


summary(glm_opens)
```

No effect - but recall that there is a lot of missing data.

# Emails versus relative compensation

Perhaps compensation is best understood relative to peers. As a quick approach, consider binning it into quintiles (and also treating NA as a separate factor level).

```{r}
CompAgg <- dat_agg %>%
  filter(!is.na(Compensation)) %>%
  group_by(email_group_filled) %>%
  summarise(
    Q1 = quantile(Compensation, probs = .2)
    , Q2 = quantile(Compensation, probs = .4)
    , Q3 = quantile(Compensation, probs = .6)
    , Q4 = quantile(Compensation, probs = .8)
    , Q5 = quantile(Compensation, probs = 1)
  )

dat_agg <- dat_agg %>%
  left_join(CompAgg, by = c('email_group_filled' = 'email_group_filled'))
```

```{r}
CompAgg %>%
  mutate_if(is.numeric, function(x) {scales::dollar(x)})
```

```{r}
dat_agg <- dat_agg %>%
  mutate(
    RelativeCompensation = case_when(
      is.na(Compensation) ~ 'NA'
      , Compensation <= Q1 ~ 'Q1'
      , Compensation <= Q2 ~ 'Q2'
      , Compensation <= Q3 ~ 'Q3'
      , Compensation <= Q4 ~ 'Q4'
      , Compensation <= Q5 ~ 'Q5'
    ) %>% factor
    , AnySignonNAs = ifelse(is.na(AnySignon), FALSE, AnySignon)
  )
```
```{r, error = FALSE, warning = FALSE, fig.width = 8, fig.height = 12}
# Check results
dat_agg %>%
  ggplot(aes(x = RelativeCompensation, y = Compensation, color = email_group_filled)) +
  geom_boxplot() +
  facet_grid(RelativeCompensation ~ email_group_filled)
```

```{r}
glm_opens2 <- dat_agg %>%
  glm(
    any_open ~ email_group_filled + same_email + send_email_status + ns(weeks_since_email, df = 4) + RelativeCompensation + AnySignonNAs
    , data = .
    , family = binomial()
  )

summary(glm_opens2)
```

No relative compensation effect either.